Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets

### 2021-07-02 main.tex
commentaires florent

%
\begin{align}
%\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q}{\bot}) & = \one\label{eq:SWVPA-weight}\\
%\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q'}{\gamma}) & = \zero 
%\mathrm{~if~} q \neq q'\nonumber\\
\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q'}{\bot}) & = \one
\mathrm{~if~} q = q' \mathrm{~and~} \zero \mathrm{~otherwise}\label{eq:SWVPA-weight}\\
\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, a\, u, 
               \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, c, p, a, q'') 
  \otimes \weight_A\bigl(\configup{q''}{\<{\call{c}}, p> \stackup \gamma}, u, 
                         \config{q'}{\gamma'}\bigr)\nonumber\\
 %
\weight_A\bigl(\config{q}{\bot}, a\, u, 
               \config{q'}{\gamma'}\bigr) & =  
  {\displaystyle\bigoplus_{q'' \in Q}} \weiei(q, a, q'') 
   \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, {\call{c}'}\, u, 
               \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{\begin{array}{c}
                          \scriptstyle q'' \in Q\\[-1ex]
                          \scriptstyle p' \in P
                          \end{array}}}
 \weic\bigl(q, {\call{c}}, p, {\call{c}'}, p', q''\bigr) 
 \otimes \weight_A\bigl(\configup{q''}{\<{\call{c}'}, p'>\stackup \<{\call{c}}, p>\stackup \gamma}, u, 
                        \config{q'}{\gamma'}\bigr)\nonumber\\[1mm]
%
\weight_A\bigl(\config{q}{\bot}, {\call{c}}\, u, 
               \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{\begin{array}{c}
                          \scriptstyle q'' \in Q\\[-1ex]
                          \scriptstyle p \in P
                          \end{array}}}
  \weiec(q, {\call{c}}, p, q'') 
  \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p}, u, 
                         \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup \gamma}, {\return{r}}\, u, 
               \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} 
  \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
  \otimes \weight_A\bigl(\config{q''}{\gamma}, u, 
                         \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, 
               \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} \weier(q, {\return{r}}, q'') 
  \otimes \weight_A\bigl(\config{q''}{\bot}, u, 
                         \config{q'}{\gamma'}\bigr)\nonumber
\end{align}
%
% and ${\call{c}}\, p\stacksep \gamma$ 
%denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol 
%and $\gamma$ is the rest of stack.

@@

%
\noindent
If the stack is not empty, let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
and let $b_\top : Q \times P \times Q \to \Phic$ be such that,
for every two states $q, q' \in Q$ 
and stack symbol $p \in P$: %and $\sigma \in \{ \bot, \top \}$, 
\begin{equation}\label{eq:btop}
  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Omega^*} 
  \textstyle
  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr)
\end{equation}
%
Intuitively, the function defined in \eqref{eq:btop}
associates to $c \in \Omegac$ 
the minimum weight of a computation of $A$
starting in state $q$ with a stack 
$\<c, p> \cdot \gamma \in \Gamma^+$ 
and ending in state $q'$ with the same stack,
such that the computation can not pop 
the pair made of $c$ and $p$ at the top of this stack,
but may only read these symbols.
Moreover, $A$ may push another pair $\< c', p'>$ %call symbols 
on the top of $\<c, p> \cdot \gamma$,
following the third case of 
in the definition \eqref{eq:SWVPA-weight} of $\weight_A$,
and may pop $\< c', p'>$ later, following the fifth case of \eqref{eq:SWVPA-weight} (return symbol). 
%However, it cannot apply one of the two last cases (return symbol and empty stack)
%when the current stack is $\gamma$.
%pop symbols in $\gamma$.
% Note that having a stack reduced to such a symbol makes impossible the application of the 
% two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
% However, it is possible to apply the two first cases 
% (internal symbol or call symbol, with a push on the top of $\top$).

@@

\noindent
The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra}, 
Figure~\ref{fig:best-update}
are well defined
since~$\Semiring$ is complete.
\florent{complete **} 
** effectively computable by hypothesis that the label theory is effective**
The algorithm performs $2.|Q|^2$ iterations until $P$ is empty, 
and each iteration has a time complexity $O(|Q|^2 . |P|)$.
\florent{detail with nb tr. and states}
That gives a time complexity $O(|Q|^4 . |P|)$. 
It can be reduced by implementing $P$ as a priority queue, 
prioritized by the value returned by $d$.
%$|Q|^3.\log(|Q|^2)$

@@

%\paragraph{Application to Automated Music Transcription.}
%\input{transcription}
%...

@@

% discussion
\noindent
Ongoing and future work include

>>>>

\florent{notation $cp$ for $\< c, p>$ ?}
%
\begin{align}
%\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q}{\bot}) & = \one\label{eq:SWVPA-weight}\\
%\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q'}{\gamma}) & = \zero 
%\mathrm{~if~} q \neq q'\nonumber\\
\weight_A\bigl(\config{q}{\bot}, \varepsilon, \config{q'}{\bot}) & = \one
\mathrm{~if~} q = q' \mathrm{~and~} \zero \mathrm{~otherwise}\label{eq:SWVPA-weight}\\
\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, a\, u, 
               \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} \weii(q, c, p, a, q'') 
  \otimes \weight_A\bigl(\configup{q''}{\<{\call{c}}, p> \stackup \gamma}, u, 
                         \config{q'}{\gamma'}\bigr)\nonumber\\
 %
\weight_A\bigl(\config{q}{\bot}, a\, u, 
               \config{q'}{\gamma'}\bigr) & =  
  {\displaystyle\bigoplus_{q'' \in Q}} \weiei(q, a, q'') 
   \otimes \weight_A\bigl(\config{q''}{\bot}, u, \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup\gamma}, {\call{c}'}\, u, 
               \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{\begin{array}{c}
                          \scriptstyle q'' \in Q\\[-1ex]
                          \scriptstyle p' \in P
                          \end{array}}}
 \weic\bigl(q, {\call{c}}, p, {\call{c}'}, p', q''\bigr) 
 \otimes \weight_A\bigl(\configup{q''}{\<{\call{c}'}, p'>\stackup \<{\call{c}}, p>\stackup \gamma}, u, 
                        \config{q'}{\gamma'}\bigr)\nonumber\\[1mm]
%
\weight_A\bigl(\config{q}{\bot}, {\call{c}}\, u, 
               \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{\begin{array}{c}
                          \scriptstyle q'' \in Q\\[-1ex]
                          \scriptstyle p \in P
                          \end{array}}}
  \weiec(q, {\call{c}}, p, q'') 
  \otimes \weight_A\bigl(\config{q''}{{\call{c}}\, p}, u, 
                         \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\configup{q}{\<{\call{c}}, p>\stackup \gamma}, {\return{r}}\, u, 
               \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} 
  \weir\bigl(q, {\call{c}}, p, {\return{r}}, q''\bigr) 
  \otimes \weight_A\bigl(\config{q''}{\gamma}, u, 
                         \config{q'}{\gamma'}\bigr)\nonumber\\
%
\weight_A\bigl(\config{q}{\bot}, {\return{r}}\, u, 
               \config{q'}{\gamma'}\bigr) & =  
 {\displaystyle\bigoplus_{q'' \in Q}} \weier(q, {\return{r}}, q'') 
  \otimes \weight_A\bigl(\config{q''}{\bot}, u, 
                         \config{q'}{\gamma'}\bigr)\nonumber
\end{align}
%
% and ${\call{c}}\, p\stacksep \gamma$ 
%denotes a stack where the pair made of ${\call{c}} \in \Omegac$ and $p \in P$ is the top symbol 
%and $\gamma$ is the rest of stack.

@@

%
\noindent
If the stack is not empty, let $\top$ be a fresh stack symbol which does not belong to $\Gamma$,
\florent{$b_\top$ : mot bien parenthèsé $c$/$r$}
and let $b_\top : Q \times P \times Q \to \Phic$ be such that,
for every two states $q, q' \in Q$ 
and stack symbol $p \in P$: %and $\sigma \in \{ \bot, \top \}$, 
\begin{equation}\label{eq:btop}
  b_\top(q, p, q') : c \mapsto \bigoplus_{s\in \Omega^*} 
  \textstyle
  \weight_A\bigl(\configup{q}{\< c, p> \stackup \top }, s, \configup{q'}{\<c, p> \stackup \top}\bigr)
\end{equation}
%
Intuitively, the function defined in \eqref{eq:btop}
associates to $c \in \Omegac$ 
the minimum weight of a computation of $A$
starting in state $q$ with a stack 
$\<c, p> \cdot \gamma \in \Gamma^+$ 
and ending in state $q'$ with the same stack,
such that the computation can not pop 
the pair made of $c$ and $p$ at the top of this stack,
but may only read these symbols.
Moreover, $A$ may push another pair $\< c', p'>$ %call symbols 
on the top of $\<c, p> \cdot \gamma$,
following the third case of 
in the definition \eqref{eq:SWVPA-weight} of $\weight_A$,
and may pop $\< c', p'>$ later, following the fifth case of \eqref{eq:SWVPA-weight} (return symbol). 
%However, it cannot apply one of the two last cases (return symbol and empty stack)
%when the current stack is $\gamma$.
%pop symbols in $\gamma$.
% Note that having a stack reduced to such a symbol makes impossible the application of the 
% two last cases in the definition of $\weight_A$ (return symbol and empty stack). 
% However, it is possible to apply the two first cases 
% (internal symbol or call symbol, with a push on the top of $\top$).

@@

\noindent
\florent{explication Fig.~\ref{fig:best-update} suivant cas de \eqref{eq:SWVPA-weight}}
The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra}, 
Figure~\ref{fig:best-update}
are well defined
since~$\Semiring$ is complete.
\florent{complete **} 
** effectively computable by hypothesis that the label theory is effective**
The algorithm performs $2.|Q|^2$ iterations until $P$ is empty, 
and each iteration has a time complexity $O(|Q|^2 . |P|)$.
\florent{detail with nb tr. and states}
That gives a time complexity $O(|Q|^4 . |P|)$. 
It can be reduced by implementing $P$ as a priority queue, 
prioritized by the value returned by $d$.
%$|Q|^3.\log(|Q|^2)$

@@

%\paragraph{Application to Automated Music Transcription.}
%\input{transcription}
%...
\florent{2 lines Application to Automated Music Transcription: 
implementation $\neq$ but same principle, on-the-fly automata construction
during best search, for efficiency.}

@@

% discussion
\noindent
Ongoing and future work include 
\florent{TODO future work}
